---
title: "Chapter 4 Exercises and Supporting Materials"
author: "J. Seawright"
date: "2024-01-04"
output:
  html_document:
    css: lab.css
    highlight: pygments
    theme: sandstone
    toc: true
    toc_float: true
---

```{r global_options, include = FALSE}
knitr::opts_chunk$set(eval = TRUE, results = FALSE)
library(tidyverse)
library(openintro)
```

## A simple regression example

What causes inequality around the world? This question has received a great deal of research attention from social scientists of all kinds, and we will not solve it in these exercises. Nevertheless, an overly simplistic approach to the question can be useful as a stylized example. Consider the inequality dataset, available on the github site:

```{r, echo=TRUE}
inequality <- read.csv("https://github.com/jnseawright/practice-of-multimethod/raw/main/data/inequality.csv")
```

We can start by briefly examining the contents of the dataset:

```{r, echo=TRUE}
summary(inequality)
```

Let's check whether democracy is connected with lower levels of economic equality, compared with dictatorships. We can do this by carrying out a bivariate regression, using the Gini coefficient (a measure of inequality) as the dependent variable and the Polity variable (a measure of democracy) as the independent
variable. Recall that Gini coefficients increase as inequality gets worse, and also that the Polity measure increases as democracy goes up.

```{r, echo=TRUE}
#install.packages(jtools)
library(jtools)
inequalitylm.1 <- lm(Gini ~ Polity, data=inequality)
summ(inequalitylm.1)
```

What do we see here? Describe and explain the results, as well as any assumptions needed in order for them to be meaningful.

One might like to incorporate control variables, obviously. The most commonly used control variable in regressions of this kind is logged GDP. We can include that in a simple multivariate regression:

```{r, echo=TRUE}
inequalitylm.2 <- lm(Gini ~ Polity + I(log(GDP)), data=inequality)
summ(inequalitylm.2)
```

What does this regression show us? What assumptions do we need in order for these results to be taken seriously? What, in fact, should we do next?

## Control Variables in a Literature
	
This problem works well as a group activity. 
	
Find three to five quantitative articles on the same outcome (for example, three articles on the causes of civil wars). Make a master list of all the control variables used in those articles, as well as which articles use each variable. For each variable, determine whether it might be a good control variable that helps eliminate confounding --- but also whether it might be a bad control variable that is either an instrument or post-treatment. How close do your articles come to including the good control variables and excluding the bad ones?

How much do the articles across a particular subject area agree about the control variables that should be used, and how much do they disagree? When they disagree, is it because they are exploring different causes, or is it for some other reason?

## Discussion Questions

Find an example of research using regression in a way that you regard as successful from your area of research. What makes this application of regression successful? What is regression used for, what assumptions are needed, and to what extent do you regard the results as credible?

This book's discussion of process tracing has drawn on Bayesian ideas, but has not gone all the way toward full formal Bayesian math. Would it be helpful for multi-method research if qualitative scholars fully formalized their research as Bayesian statistical problems, or would that get in the way of the qualitative scholarship? 